HydroFlow: A New Dataflow Programming Language

### 2022-10-31 paper.tex
i continue to dislike this text

>>>>

\section{An overview of the HydroFlow language}
\label{sec:hydroflow-overview}

\subsection{Hydroflow as a dataflow library}

At its most basic, hydroflow functions as a traditional dataflow
library; it applies a sequence of operators to some stream,
transforming a stream of inputs into a stream of outputs. As with
modern stream-oriented libraries, Hydroflow's notion of streams is
highly flexible; operators, which may be built-in or user-supplied,
are able to consume and produce any number of streams (as inputs and
outputs respectively), and may maintain internal state.

\subsubsection{Defining Hydroflow operators}

Operationally, HydroFlow operators are defined as ordinary Rust
functions which take any number of arguments, each of which
corresponds to a single element from the operator's input stream(s).

...I actually don't remember the details of this part, time to read the hydroflow book.

Our basic goals for this section are to say:

\subsubsection{Assembling a Hydroflow Graph}

With operators in hand, users may now assemble a hydroflow graph.  To
do this, we have supplied a convenient high-level graph assembly
syntax reminiscent of XXX, whose formal definition can be found in
figure \ref{fig:hydroflow-pipeline-syntax}. With this syntax, users
explicitly create instances of previously-defined operators, and
``wire'' them together via the "->" notation.  For operators with
multiple input or output streams, users can index operators via the
"[n]op" (for inputs) or "op[n]" (for outputs) syntax.  Our
graph-assembly compiler ensures that all inputs and outputs to
operators are successfully connected, and that no output (or input) is
used twice.

\subsubsection{Running a hydroflow program}

I'm not sure there's a ton to say about this, but here's a stub.

\subsection{Hydroflow as a stream-oriented programming language}

When viewed as a language, Hydroflow's power and promise above
traditional dataflow libraries reveals itself. Most interesting are
two key points: hydroflow programs as transformers for streams of
lattice points, and Hydroflow's ability to reason about _properties_
of these streams.

\subsubsection{Streams of lattice points}

In Hydroflow, a _lattice_\footnote{as in lattice order, from order
theory} is a special class of datastructure equipped with a partial
order ⊑ and _join_q (written ⊔) and _meet_ (written ⊓) functions.
Both join and meet must be associative, commutative, and idempotent;
additionally, join must inflate elements along ⊑, while meet must
deflate them. That is: $∀ a,b: a ⊑ a ⊔ b ∧ b ⊑ a ⊔ b$ (for join) and
$∀ a,b: a ⊓ b ⊑ a ∧ a ⊓ b ⊑ b$ (for meet).  Rather than operate on
streams of arbitrary data types, hydroflow programs operate on streams
of _Lattice Points_, or elements of lattices, drawn from any number of
user-defined Lattices.

Our choice to use lattices (and lattice points) provides key
information to the hydroflow compiler when determining optimizations;
which, of course, it does not take advantage of in any way whatsoever,
so I guess this section needs to tie into the story about informing a
higher-level analysis about lattice stuff.

While our choice to use lattices alone might seem more restrictive
than traditional dataflow engines, the reality is in fact the
opposite; one can always build a lattice from an arbitrary datatype.
Rather, our choice to privilege streams of lattice points provides the
opportunity to the programmer to communicate certain datastructure
features to the hydroflow compiler; should those features be absent,
one can always take the free lattice over a given datatype. This is
achieved by declaring all elements of that datatype to be unordered by
⊑, and introducing an arbitrary ⊤ and ⊥ elements to serve as the
supremum and infinimum of the lattice respectively.  One can then
define $a ⊔ b ≜ (a == b ? a : ⊤)$ and $a ⊓ b ≜ (a == b ? a : ⊥)$,
satisfying the requirements for both ⊔ and ⊓ operators.

\subsubsection{Properties of streams}
